### 2012-06-30 doc/semester/report.tex
blurb

\documentclass[preprint,9pt]{sigplanconf}

\input{dot_macros}

    \usepackage[pdftex,
                hyperindex,
                plainpages=false,
                breaklinks,
                colorlinks,
                citecolor=black,
                filecolor=black,
                linkcolor=black,
                pagecolor=black,
                urlcolor=black]{hyperref}
    \usepackage[pdftex]{graphicx}
    \DeclareGraphicsExtensions{.jpg,.pdf}
    \pdfcatalog {
        /PageMode (/UseNone)
    }
    \usepackage{thumbpdf}
    \usepackage[pdftex]{color}

\begin{document}

\conferenceinfo{EDIC}{July 2012, EPFL.}
\copyrightyear{2012}
\copyrightdata{[to be supplied]}

\titlebanner{Dependent Object Types}                           % These are ignored unless
\preprintfooter{Spring 2012 Semester Project -- Nada Amin}     % 'preprint' option specified.

\title{Dependent Object Types}
\subtitle{Spring 2012 Semester Project}

\authorinfo{Nada Amin}
           {EPFL}
           {nada.amin@epfl.ch}

\maketitle

\begin{abstract}
In my semester project, I have studied a new proposed type-theoretic
foundation of Scala and languages like it: the Dependent Object Types
calculus (DOT). DOT models Scala's path-dependent types and abstract
type members, as well as its mixture of nominal and structural typing
through the use of refinement types. My ultimate goal was to prove DOT
sound, but instead, I found lots of counterexamples to soundness, and
have explored patches to the calculus.
\end{abstract}

\section{Introduction}

% TODO

\section{Counterexamples}

\subsection{Subtyping Transitivity and Preservation}

Some subtyping transitivity is essential for soundness. Indeed, I show
how to construct a counterexample to preservation from any
counterexample to subtyping transitivity where the three involved
types $S, T, U$ are expressible within a universe $u$ but not
necessarily realizable:

\begin{align*}
\mlnew u \ldots {
\ \mlapp{\abs x \Top x}{
\ \ \mlapp{\abs f {S \tfun U} f}{
\ \ \ \mlapp{\abs f {S \tfun T} f}{
\ \ \ \ \mlapp{\abs f {S \tfun S} f}{
\ \ \  \ \ \abs x S x}}}}}
\end{align*}

The idea is to start with a function from $S \tfun S$ and cast it
successively to $S \tfun T$ then $S \tfun U$. To typecheck the
expression initially, we need to check $S \sub T$ and $T \sub
U$. After some reduction steps, the first few casts vanish, and the
reduced expression casts directly from $S \tfun S$ to $S \tfun U$, so
we need to check $S \sub U$. Thus, we need subtyping transitivity: $S
\sub T$ and $T \sub U$ implies $S \sub U$.

\subsection{Non-Expanding Types and Subtyping Transitivity}

The well-formedness rule $\textsc{tsel-wf$_2$}$ allow a type selection
$p.L$ to refer back to itself in its upper bound. This is useful for
expressing recursive class types and F-bounded abstract types.

Consider the simplest possible type that refers to itself: $T = \Top
\refine z {\Ldecl L \Bot z.L}$. This type is well-formed but
non-expanding.

Non-expanding types are problematic for subtyping transitivity.

\acks

I thank Adriaan Moors and Martin Odersky for sharing previous work,
fruitful discussions and guidance.

%\bibliographystyle{abbrvnat}
%\bibliography{bib}

\end{document}