### 2012-06-30 doc/semester/report.tex
term-mem restriction

\documentclass[preprint,9pt]{sigplanconf}

\input{dot_macros}

    \usepackage[pdftex,
                hyperindex,
                plainpages=false,
                breaklinks,
                colorlinks,
                citecolor=black,
                filecolor=black,
                linkcolor=black,
                pagecolor=black,
                urlcolor=black]{hyperref}
    \usepackage[pdftex]{graphicx}
    \DeclareGraphicsExtensions{.jpg,.pdf}
    \pdfcatalog {
        /PageMode (/UseNone)
    }
    \usepackage{thumbpdf}
    \usepackage[pdftex]{color}

\begin{document}

\conferenceinfo{EDIC}{July 2012, EPFL.}
\copyrightyear{2012}
\copyrightdata{[to be supplied]}

\titlebanner{Dependent Object Types}                           % These are ignored unless
\preprintfooter{Spring 2012 Semester Project -- Nada Amin}     % 'preprint' option specified.

\title{Dependent Object Types}
\subtitle{Spring 2012 Semester Project}

\authorinfo{Nada Amin}
           {EPFL}
           {nada.amin@epfl.ch}

\maketitle

\begin{abstract}
In my semester project, I have studied a new proposed type-theoretic
foundation of Scala and languages like it: the Dependent Object Types
calculus (DOT). DOT models Scala's path-dependent types and abstract
type members, as well as its mixture of nominal and structural typing
through the use of refinement types. My ultimate goal was to prove DOT
sound, but instead, I found lots of counterexamples to soundness, and
have explored patches to the calculus.
\end{abstract}

\section{Introduction}

% TODO

\section{Counterexamples}

\subsection{Subtyping Transitivity and Preservation}\label{subpres}

Some subtyping transitivity is essential for soundness. Indeed, I show
how to construct a counterexample to preservation from any
counterexample to subtyping transitivity where the three involved
types $S, T, U$ are expressible within a realizable universe, though
the types themselves don't need to be realizable:

\begin{align*}
&\mlnew u \ldots {
\ \mlapp{\abs x \Top x}{
\ \ \mlapp{\abs f {S \tfun U} f}{
\ \ \ \mlapp{\abs f {S \tfun T} f}{
\ \ \ \ \mlapp{\abs f {S \tfun S} f}{
\ \ \  \ \ \abs x S x}}}}}
\end{align*}

The idea is to start with a function from $S \tfun S$ and cast it
successively to $S \tfun T$ then $S \tfun U$. To typecheck the
expression initially, we need to check $S \sub T$ and $T \sub
U$. After some reduction steps, the first few casts vanish, and the
reduced expression casts directly from $S \tfun S$ to $S \tfun U$, so
we need to check $S \sub U$. Thus, we need subtyping transitivity: $S
\sub T$ and $T \sub U$ implies $S \sub U$.

\subsection{Non-Expanding Types and Subtyping Transitivity}\label{nonexp}

The well-formedness rule $\textsc{tsel-wf$_2$}$ allow a type selection
$p.L$ to refer back to itself in its upper bound. This is useful for
expressing recursive class types and F-bounded abstract types.

Consider the simplest possible type that refers to itself: $p \typ T =
\Top \refine z {\Ldecl L \Bot z.L}$. Now, $p.L$ is well-formed but
non-expanding. Indeed, the only expansion rule that applies to $p.L$
is $\textsc{tsel-$\expand$}$, but it requires expanding the upper
bound... $p.L$! Thus, there is no finite derivation for the expansion
of $p.L$.

Non-expanding types are problematic for subtyping
transitivity. Consider: $p.L \sub \Top$ by \textsc{$\sub$-$\Top$} and
$\Top \sub \Top \refine z {}$ by \textsc{$\sub$-rfn}. By transitivity
on $\Top$, we expect $p.L \sub \Top \refine z {}$, but we cannot infer
this from the rules. The only rules that apply are \textsc{$\sub$-rfn}
and \textsc{tsel-$\sub$}. \textsc{$\sub$-rfn} gets stuck expanding
$p.L$. \textsc{tsel-$\sub$} gets stuck because the conclusion is also
a premise as $p \ni \Ldecl L \Bot p.L$.

For this counterexample, it seems enough to just add transitivity on
$\Top$ as a built-in rule. But this doesn't solve the general issue
that non-expanding types don't play well with subtyping. The problem
is deep, as it's possible to construct a realizable universe with
three non-trivial types that fail subtyping transitivity.

Consider an environment where $u$ is bound to:
\begin{align*}
\Top & \mlrefine u {\\
&\ \Ldecl {\mathit{Bad}} {\Bot} {u.\mathit{Bad}}\\
&\ \Ldecl {\mathit{Good}} {\Top \refine z {\Ldecl L \Bot \Top}} {\Top \refine z {\Ldecl L \Bot \Top}}\\
&\ \Ldecl {\mathit{Lower}} {u.\mathit{Bad} \tand u.\mathit{Good}} {u.\mathit{Good}}\\
&\ \Ldecl {\mathit{Upper}} {u.\mathit{Good}} {u.\mathit{Bad} \tor u.\mathit{Good}}\\
&\ \Ldecl X {u.\mathit{Lower}} {u.\mathit{Upper}}\\
}&
\end{align*}
Note that each lower bound is a subtype of its upper bound, so $u$ is
realizable. Hence, if we find a counterexample to subtyping
transitivity with types expressible within $u$, we can apply the trick
from section~\ref{subpres} to create a counterexample to preservation.

Indeed, here is such a counterexample to subtyping transitivity:
\begin{align*}
S &= u.\mathit{Bad} \tand u.\mathit{Good}\\
T &= u.\mathit{Lower}\\
U &= u.X \refine z {\Ldecl L \Bot \Top}
\end{align*}
We have $S <: T$ and $T <: U$, but we cannot derive $S <: U$ because
$S$ doesn't expand.

\subsection{Narrowing}\label{narrowing}

\subsubsection{Functions as Objects}\label{fun}

The original DOT calculus has both object and function types. It
doesn't have built-in methods, but they can be expressed as functional
labels. On the other hand, Scala doesn't have a built-in function
type. Instead, it has both methods and fields, and encodes functions
as objects with an apply method. Here, I show that the DOT model as
originally designed is problematic.

A concrete object can be a subtype of a function type without a
function ever being defined. Consider:
\begin{align*}
&\mlnew u {\Top \refine z {\Ldecl C {\Top \tfun \Top} {\Top \tfun \Top}} \ldefs{}} {
\mlnew f {u.C \ldefs{}} {
\ldots
}}
\end{align*}

Now, $f$ is a subtype of $\Top \tfun \Top$, but $\app f {(\abs x \Top
  x)}$ is stuck (and, rightfully, doesn't typecheck). But we can use
narrowing to create a counterexample to preservation: $\app {(\abs g
  {\Top \tfun \Top} {g (\abs x \Top x)})} f$.

\subsubsection{\texorpdfstring{\textsc{term-$\ni$}}{Term-Mem} Restriction}\label{term_mem}

There are two membership ($t \ni D$) rules: one for when the term $t$
is a path, and one for an arbitrary term $t$. For paths, we can
substitute the self-references in the declarations, but we cannot do
so for arbitrary terms as the resulting types wouldn't be well-formed
syntactically. Hence, the $\textsc{term-$\ni$}$ has the restriction
that self-ocurrences are not allowed. Here is a counterexample related
to this restriction.

Let $X$ be a shorthand for the type:
\begin{align*}
\Top & \mlrefine z {\\
&\ \Ldecl {L_a} \Top \Top\\
&\ \ldecl l {z.L_a}\\
}&
\end{align*}

Let $Y$ be a shorthand for the type:
\begin{align*}
\Top & \mlrefine z {\\
&\ \ldecl l \Top\\
}&
\end{align*}

Now, consider the term
\begin{align*}
&\mlnew u {X \ldefs{ l = u }}{
(\app {(\abs y {\Top \tfun Y} {\app y u})} {(\abs d \Top {(\app {(\abs x X x)} u)})}).l
}
\end{align*}

The term type-checks because the $t=(\app {(\abs y {\Top \tfun Y}
  {\app y u})} {(\abs d \Top {(\app {(\abs x X x)} u)})})$ has type
$Y$, so we can apply $\textsc{term-$\ni$}$ for $l$. However, the term
$t$ eventually steps to $(\app {(\abs x X x)} u)$ which has type $X$,
so we cannot apply $\textsc{term-$\ni$}$ for $l$ because of the
self-reference ($z.L_a$).

\subsubsection{Expansion Lost}\label{narrowing_exp}

\subsubsection{Well-Formedness Lost}\label{narrowing_wf}

\subsection{Path Equality}\label{patheq}

\acks

I thank Adriaan Moors and Martin Odersky for sharing previous work,
fruitful discussions and guidance.

%\bibliographystyle{abbrvnat}
%\bibliography{bib}

\end{document}